\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2016-07-21 redaction/dhp_draft_july2016.tex
Modif intro

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad valorem tax equivalent. However, many common empirical facts support the existence of additive costs. This paper provides results on the size and composition of transport costs over time, explicitly distinguishing between its multiplicative and additive components, using sectorial cif-fob prices over 1974-2013 taken from US imports data. First, we estimate the two components of transport costs, by transport mode (air or ocean). We thus find that ocean transport costs decompose in an ad-valorem tax equivalent of mean value 3.22\% and a per-unit cost that amounts to 2.85\% of the mean fob price over the period. These values are respectively equal to 2.5\% and 1.9\% in air shipping. Second, we provide a quantitative assessment what standard international trade models lose by skipping additive transport costs. Quantitatively, the omission of the additive term leads to overestimate the iceberg component by roughly a factor 2. On average over the whole period, biased estimate for iceberg is 5\% for Air and 6\% for Vessel, while unbiased estimate is respectively 2.5\% and 3.2\%. Third, the time dimension of our data allows us to characterize the evolution of transport costs. After excluding the composition effects, we obtain that transport costs \textit{per se} have roughly decreased by 40\% over the period 1974-2012. While this order of magnitude applies for both the additive and multiplicative components of vessel transport costs, air transport costs exhibit more contrasted trends.
\end{abstract}

>>>>

\begin{abstract}
Following \cite{samuelson1954}, standard models of international trade have usually relied on modelling trade costs as an ad valorem tax equivalent. However, many common empirical facts support the existence of additive costs. This paper measures the size of additive costs in transport, using SITC 3 and 4 digit cif-fob unit values over 1974-2013 taken from US imports data. We estimate the two components of transport costs, by transport mode (air or ocean). We find that additive costs are 2.85\% of fob unit values for ocean transport costs (and ad-valorem ones 3.22\%).  These values are respectively equal to 1.9\% and 2.5\% for air transport costs. We show that taking additive costs into account improves the fit of the modelling of transport costs. The time dimension of our data allows us to characterize the evolution of transport costs. After correcting for composition effects, we obtain that all types of transport costs have been constant from 1974 to 1984 and then  decreased by 40\% over the period 1984-2013. Most of the early decline in air transport costs can be explained by the ad-valorem component. By contrast, this component nearly doubled in the 2000s. 
\end{abstract}
